Theories of deliberated choice

### 2024-01-23 simple.tex
Cleaned some ¬HAI

  Let us now prove that \ref{it:condNotHAI} implies \ref{it:condNotHAIInf}.
  Consider any $\hist \in \allhists, i_1 \in I$.
  Define $J = \set{i_2 \in I \suchthat \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)}$.
  Assume that $\forall i_2 \in J: \pi(i_2) \neq \emptyset$ (otherwise, the first disjunct of \ref{it:condNotHAIInf} holds), and let us show that $\card\set{i_2 \in J \suchthat \pi(i_1) \neq \pi(i_2)} = \infty$.
  From $\forall i_2 \in J: \pi(i_2) \neq \emptyset$, we obtain $\pi(i_1) \neq \emptyset$, thus, $\exists a_{i_1} \in \allargs \suchthat a_{i_1} \ileadstost[i_1] \pi(i_1)$.
  Apply \ref{it:condNotHAI} to obtain some $i_2 \in J$ such that $\pi(i_1) \neq \pi(i_2)$.
  Also, by our assumption, $\pi(i_2) \neq \emptyset$, thus, $\exists a_{i_2} \in \allargs \suchthat a_{i_2} \ileadstost[i_2] \pi(i_2)$.
  Consider the path $\histp = (\hist, a_{i_1}, a_{i_2})$. We have $\histp \ileadsto[i_1] \pi(i_1) \neq \histp \ileadsto[i_2] \pi(i_2)$, thus ${\ileadsto[i_1]}(\histp) \neq {\ileadsto[i_2]}(\histp)$. 
  Apply \ref{it:condNotHAI} on $\histp, i_1$. We obtain some $i_3 \in I$ such that $\pi(i_1) \neq \pi(i_3)$ and $\forall \histpp \wsub \histp: {\ileadsto[i_1]}(\histpp) = {\ileadsto[i_3]}(\histpp)$.
  In particular, ${\ileadsto[i_1]}(\histp) = {\ileadsto[i_3]}(\histp)$, thus, $i_3 \neq i_2$.
  Also, $i_3 \in J$.
  Iterating this process, we obtain an infinite sequence of distinct individuals in $J$.
  In other words, we have shown that if some $i_2 \in J, \hist^{(2)} \in \allhists$ with $\pi(i_2) \neq \emptyset$ and $i_1$ and $i_2$ agree everywhere on $\hist^{(2)}$, then there exists $i_3 \in J, \hist^{(3)} \in \allhists$ with $\pi(i_3) \neq \emptyset$ and $\hist^{(2)} \ssub \hist^{(3)}$ and $i_1$ and $i_2$ disagree somewhere on $\hist^{(3)}$ and $i_1$ and $i_2$ agree everywhere on $\hist^{(3)}$. (Should check this summary.)
  Could also use the following probably simpler proof. Consider any finite subset of $j$ individuals, build a $H'$ with all the test paths containing decisive arguments, namely $\hist^{(2)} = (\hist, a_1, a_2)$, $\hist^{(3)} = (\hist^{(2)}, a_1, a_3)$, …, and show that there’s another $j$ in fact.
\end{proof}
\begin{example}
  TODO find example where Per \cref{th:caractNotIndic}, there is no claim that holds and is anonymous and indicative.

  % Given $\ar \in \allargs$ and $p, p' \in P$, define ${\ileadsto[i^{(\ar, p, p')}]}$ as $\hist \ileadsto[i^{(\ar, p, p')}] p$ iff $\ar \in \histend$ and $\hist \ileadsto[i^{(\ar, p, p')}] p'$ otherwise.
  % It follows that $\pi(i^{(\ar, p, p')}) = p$.

  % Pick any two $p \neq p' \in P$ and consider infinitely many arguments $\allargs = \set{\ar_1, …, }$ and infinitely many individuals $I = \set{i^{(\ar, p, p')} \suchthat \ar \in \allargs} \cup \set{i^{(\ar, p', p)} \suchthat \ar \in \allargs}$.
  % Per \cref{th:caractNotIndic}, there is no claim that holds and is anonymous and indicative:
  % consider any $\hist, i_1$, if $\ar_1$. NOPE, that’s wrong, I think.

  % Define ${\ileadsto[i^{(\ar, \phi)}]}$ as $\hist \ileadsto[i^{(\ar, \phi)}] \phi$ iff $\hist_0 = \ar$ and $\hist \ileadsto[i^{(\ar, \phi)}] ¬\phi$ otherwise.
  % It follows that $\pi(i^{(\ar, \phi)}) = \emptyset$.

  The condition $\pi(i_1) \neq \emptyset$ in \ref{it:condNotHAI} is required, as seen by this example where \ref{it:condNotHAI} is satisfied, even though
  $[\forall \hist \in \allhists, i_1 \in I: \exists i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)]$ is not.
We could have \ref{it:condNotHAI} satisfied, but not the part with $\pi(i_1) \neq \emptyset$.
  Consider $i_1$ with empty DJ on some path has only individuals with identical behavior on the path that have an empty DJ (though for normal $i$’s, the condition is satisfied, thus there are collisions, meaning some individuals with different DJ having identical behavior on the path).
  TODO check.
\end{example}
\begin{remark}
  Given $J \wsub I$, say that a set of paths discriminates $J$ iff some function associating each path to an answer is compatible with some $j \in J$ (meaning that $f \wsub {\ileadsto[j]}$) and is compatible with no $i \notin J$.
  The condition says that there must be some $p$ and some increasing set of paths $H$ such that $H$ discriminates $J = \pi^{-1}(p)$.
\end{remark}
\begin{remark}
  For failing the condition, I need an infinity of $j$’s (and at least one $i$) who agree on any branch of the history tree? (Meaning relations $\ileadsto[j]$, we don’t care about individuals.)
  No: on any finite branch in the tree, I need that remains one $j$ and one $i$ such that infinite prolongation of the branch will never discriminate them.

>>>>

  Let us now prove that \ref{it:condNotHAI} implies \ref{it:condNotHAIInf}.
  Consider any $\hist \in \allhists, i_1 \in I$.
  Define $J = \set{i_2 \in I \suchthat \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)}$.
  Assume that $\forall i_2 \in J: \pi(i_2) \neq \emptyset$ (otherwise, the first disjunct of \ref{it:condNotHAIInf} holds), and let us show that $\card\set{i_2 \in J \suchthat \pi(i_1) \neq \pi(i_2)} = \infty$.
  From $\forall i_2 \in J: \pi(i_2) \neq \emptyset$, we obtain $\pi(i_1) \neq \emptyset$, thus, $\exists a_{i_1} \in \allargs \suchthat a_{i_1} \ileadstost[i_1] \pi(i_1)$.
  Apply \ref{it:condNotHAI} to obtain some $i_2 \in J$ such that $\pi(i_1) \neq \pi(i_2)$.
  Also, by our assumption, $\pi(i_2) \neq \emptyset$, thus, $\exists a_{i_2} \in \allargs \suchthat a_{i_2} \ileadstost[i_2] \pi(i_2)$.
  Consider the path $\histp = (\hist, a_{i_1}, a_{i_2})$. We have $\histp \ileadsto[i_1] \pi(i_1) \neq \histp \ileadsto[i_2] \pi(i_2)$, thus ${\ileadsto[i_1]}(\histp) \neq {\ileadsto[i_2]}(\histp)$. 
  Apply \ref{it:condNotHAI} on $\histp, i_1$. We obtain some $i_3 \in I$ such that $\pi(i_1) \neq \pi(i_3)$ and $\forall \histpp \wsub \histp: {\ileadsto[i_1]}(\histpp) = {\ileadsto[i_3]}(\histpp)$.
  In particular, ${\ileadsto[i_1]}(\histp) = {\ileadsto[i_3]}(\histp)$, thus, $i_3 \neq i_2$.
  Also, $i_3 \in J$.
  Iterating this process, we obtain an infinite sequence of distinct individuals in $J$.
  In other words, we have shown that if some $i_2 \in J, \hist^{(2)} \in \allhists$ with $\pi(i_2) \neq \emptyset$ and $i_1$ and $i_2$ agree everywhere on $\hist^{(2)}$, then there exists $i_3 \in J, \hist^{(3)} \in \allhists$ with $\pi(i_3) \neq \emptyset$ and $\hist^{(2)} \ssub \hist^{(3)}$ and $i_1$ and $i_2$ disagree somewhere on $\hist^{(3)}$ and $i_1$ and $i_2$ agree everywhere on $\hist^{(3)}$. (Should check this summary.)
  Could also use the following probably simpler proof. Consider any finite subset of $j$ individuals, build a $H'$ with all the test paths containing decisive arguments, namely $\hist^{(2)} = (\hist, a_1, a_2)$, $\hist^{(3)} = (\hist^{(2)}, a_1, a_3)$, …, and show that there’s another $j$ in fact.
\end{proof}
\begin{example}[no HAI]
  Given any infinite set $\allargs$, define $I = \set{f \in P^{\allhists} \suchthat \pi(f) \neq \emptyset}$ and $\forall i \in I: {\ileadsto} = i$.
  Considering any $\hist \in \finite(\allhists)$ and ${\ileadsto} \in I$, pick another relation that coincides on $\hist$ but with another DJ than ${\ileadsto}$. This is possible because $\hist$ is finite, thus there is an argument that is not involved in any of its elements, which can serve as decisive one.
  Per \cref{th:caractNotIndic}, there is no claim that holds and is anonymous and indicative.

  If using empty DJs, it’s easier. Define $.* \ileadsto[i_1] \phi$, $\emptyset \ileadsto[i_2] \phi$, $a.* \ileadsto[i_2] \phi$, $b.* \ileadsto[i_2] ¬\phi$, $\emptyset \ileadsto[i_3] \phi$, $a.* \ileadsto[i_3] ¬\phi$, $b.* \ileadsto[i_3] \phi$. 
  Then $\pi(i_1) = \phi$ and $\pi(i_2) = \pi(i_3) = \emptyset$.
  Also, for any path, if it is only $\emptyset$, everybody agree; if it starts with $a$ then $i_1$ and $i_2$ agree on it; if it starts with $b$ then $i_1$ and $i_3$ agree on it.

  The condition $\pi(i_1) \neq \emptyset$ in \ref{it:condNotHAI} is required, as seen by this example where \ref{it:condNotHAI} is satisfied, even though
  $[\forall \hist \in \allhists, i_1 \in I: \exists i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)]$ is not.
  TODO reuse the example here above to satisfy the condition, but add some $i_1$ with empty DJ who on some path has only individuals with identical behavior on the path that have an empty DJ.
\end{example}
\begin{remark}
  Given $J \wsub I$, say that a set of paths discriminates $J$ iff some function associating each path to an answer is compatible with some $j \in J$ (meaning that $f \wsub {\ileadsto[j]}$) and is compatible with no $i \notin J$.
  The condition says that there must be some $p$ and some increasing set of paths $H$ such that $H$ discriminates $J = \pi^{-1}(p)$.
\end{remark}
\begin{remark}
  For failing the condition, I need an infinity of $j$’s (and at least one $i$) who agree on any branch of the history tree? (Meaning relations $\ileadsto[j]$, we don’t care about individuals.)
  No: on any finite branch in the tree, I need that remains one $j$ and one $i$ such that infinite prolongation of the branch will never discriminate them.
